Query Optimization for Dynamic Imputation

### 2017-03-02 paper/examples.tex

Before she can perform her queries, the epidemiologist must develop a strategy
for handling missing values. She currently has two options:
1) she could drop records with missing values in relevant fields,
2) she could use a standard data imputation package on all of her data. Both of these
approaches have significant drawbacks. For the query pictured in~\Cref{fig:example-query}, 
(1) drops 1492 potentially relevant tuples,
while her previous experiences with (2) have proven to be expensive. The epidemiologist needs a more complete picture
of the data, so (1) is insufficient, and she is under time pressure, so (2)
is infeasible.

@@

\Cref{fig:quality-plan} shows a quality-optimized plan that uses the impute operation $\mu$, which employs an iterative regression model to fill in missing values rather than dropping tuples.
It waits to impute \verb|demo.income| until after the final join has taken place, but other imputations take place earlier on in the plan, some before filtering and join operations.
Imputations are placed to maximize ImputeDB's estimate of the quality of the overall results.
Meanwhile, \Cref{fig:fast-plan}'s performance-optimized plan uses the drop operation $\delta$ instead of filling in missing values.
However, it is a significant improvement over dropping all tuples with missing data in any field, as it only drops tuples with missing values in attributes which are referenced by the rest of the plan. In both cases, only performing
imputation on the necessary data yields a query that is much faster
than performing imputation on the whole dataset and then executing
the query.

@@

\subsection{\ProjectName{} Workflow}
Now that the epidemiologist has \ProjectName{}, she can explore
the dataset using SQL. For each query, she repeatedly chooses a value
for $\alpha$ and runs the query until she is unsatisfied with the runtime.
Rather than commit \emph{a priori} to a runtime bound for her query,
which may be unsatisfiable, this iterative approach gives her 
immediate feedback.

>>>>

Before she can perform her queries, the epidemiologist must develop a strategy
for handling missing values. She currently has two options:
(1) she could drop records that have missing values in relevant fields,
(2) she could use a traditional imputation package on her entire dataset. Both of these
approaches have significant drawbacks. For the query pictured in~\Cref{fig:example-query}, 
(1) drops 1492 potentially relevant tuples,
while from her experience, (2) has proven to take too long. The epidemiologist needs a more complete picture
of the data, so (1) is insufficient, and for this quick exploratory analysis, (2)
is infeasible.

@@

\Cref{fig:quality-plan} shows a quality-optimized plan that uses the impute operation $\mu$,
which employs a reference imputation strategy to fill in missing values rather than dropping tuples.
It waits to impute \verb|demo.income| until after the final join has taken place, though other imputations take place earlier on in the plan, some before filtering and join operations.
Imputations are placed to maximize ImputeDB's estimate of the quality of the overall results.
Meanwhile, \Cref{fig:fast-plan}'s performance-optimized plan uses the drop operation $\delta$ instead of filling in missing values.
However, it is a significant improvement over dropping all tuples with missing data in any field, as it only drops tuples with missing values in attributes which are referenced by the rest of the plan. In both cases, only performing
imputation on the necessary data yields a query that is much faster
than performing imputation on the whole dataset and then executing
the query.

@@

\subsection{\ProjectName{} Workflow}
Now that the epidemiologist has \ProjectName{}, she can explore
the dataset using SQL. She begins by choosing a value for $\alpha$ and can adjust it up or
down until she is satisfied with her query runtime. This iterative approach gives her 
immediate feedback.
